MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2018-02-01 Pattern Recognition R2/article/article.tex
Added paper documents from dropbox

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the instances within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as this proposal's overall superior performance against other multi-instance SVM models. 
\end{abstract}

@@

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as webmining and text categorization {\citep{Andrews2002}}, image classification and retreival \citep{qi2007,cao2017weakly}, human action recognition \citep{Yi2016}, medical diagnosis and imaging {\citep{Herrera2016}}, and bioinformatics {\citep{Dietterich1997}} (drug activity prediction and protein identification).

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a novel bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without assuming the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

\iitem Deriving the dual multi-instance SVM problem, with the Karush-Kuhn-Tucker necessary and sufficient conditions for optimality. The dual is maximized with respect to the Lagrange multipliers and provides insightful information about the resulting sparse model. The dual formulation is kernelized with a Gaussian radial basis function, which calculates the distances between bag representatives.
\newpage
\iitem Devising a unique bag-representative selection method that makes no presumptions about the underlying distributions of the instances within each bag, while maintaining the default MI assumption. This approach eliminates the issue of class imbalance caused by techniques such as flattening or subsetting positive instances from each bag. The key feature of MIRSVM is its ability to identify instances (support vectors) within positive and negative bags that highly impact the model.
\end{itemize}

@@

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data~\citep{2017-KNOSYS-MIMV}. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. MIRSVM, is an iterative method that uses SVMs to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities, performance, and sparser models.

@@

SimpleMI achieves the highest rank and competitive run times because, rather than use the instances in each bag to train a model, it takes the mean value of the instances in a bag and uses that for training. Even though SimpleMI has fast run-times, its performance over the previous metrics has been shown to be random and not as effective as the bag-level methods.

@@

The experimental study showed the better performance of MIRSVM compared with state-of-the-art multi-instance support vector machines, traditional multi-instance learners, as well as ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners, such as miGraph, Bagging, and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. Our proposal, MIRSVM, performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.

\section*{Acknowledgment}
This research was supported by the Spanish Ministry of Economy and Competitiveness, project TIN2014-55252-P, and by FEDER funds.

>>>>

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the instances within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as MIRSVM's overall superior performance against other multi-instance SVM models, having an average accuracy of $82.6\%$, which is $2.5\%$ better than the best performing state-of-the-art MI classifier.
\end{abstract}

@@

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as webmining and text categorization {\citep{Andrews2002}}, image classification and retreival \citep{qi2007,cao2017weakly}, human action recognition \citep{Yi2016}, medical diagnosis and imaging {\citep{Herrera2016}}, and bioinformatics {\citep{Dietterich1997}} (drug activity prediction and protein identification). In the case of drug activity prediction, MIL allows molecules to be represented as related instances contained in a bag rather than single instances with their own labels.

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a novel bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without making specific assumptions about the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

\iitem Deriving the dual multi-instance SVM problem, with the Karush-Kuhn-Tucker necessary and sufficient conditions for optimality. The dual is maximized with respect to the Lagrange multipliers and provides insightful information about the resulting sparse model. The dual formulation is kernelized with a Gaussian radial basis function, which calculates the distances between bag representatives.

\iitem Devising a unique bag-representative selection method that makes no presumptions about the underlying distributions of the instances within each bag, while maintaining the default MI assumption. This approach eliminates the issue of class imbalance caused by techniques such as flattening or subsetting positive instances from each bag. The key feature of MIRSVM is its ability to identify instances (support vectors) within positive and negative bags that highly impact the model.
\end{itemize}

@@

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data~\citep{2017-KNOSYS-MIMV}. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. Our prososed contribution, MIRSVM, is an iterative method that uses SVMs to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities, performance, and sparser models.

@@

SimpleMI achieves the highest rank and competitive run times because, rather than using the instances in each bag to train a model, it takes the mean value of the instances in a bag and uses that for training. Even though SimpleMI has fast run-times, its performance over the previous metrics has been shown to be random and not as effective as the bag-level methods.

@@

The experimental study showed the better performance of MIRSVM compared with state-of-the-art multi-instance support vector machines, traditional multi-instance learners, as well as ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners, such as miGraph, Bagging, and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. This was especially reflected in the meta-ranks comparison, shown in Table~\ref{tab:metarank}, over Accuracy, Precision, Recall, Cohen's Kappa, AUC, and run time, where instance-level learners were ranked last, whereas bag-level and ensemble learners were better in rank, and MIRSVM ranked first. Our proposal, MIRSVM, performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.

MIRSVM could be further improved by invesitgating the following approaches. Rather than using a QP solver at each iteration of MIRSVM, a different optimal solver could be experimented with. Due to the run time implications of using a QP solver, using a different solver might improve MIRSVM's scalability. Another area to investigate would be implementing MIRSVM in high-performance distributed environments. A third concept to explore would be testing MIRSVM on highly imbalanced multi-instance datasets and observe its performance. MIRSVM could also be extended to handle multi-class multi-instance problems by using traditional methods such as one-versus-one, where $c(c - 1)/2$ classifiers are trained for a $c$-class problem, and one-versus-all, where a single classifier is trained per class~\citep{bishop2007pattern}.

\section*{Acknowledgment}
This research was supported by the Spanish Ministry of Economy and Competitiveness, project TIN2017-83445-P and by FEDER funds.

### 2018-02-01 Pattern Recognition R2/article_with_markings/article.tex
Added paper documents from dropbox

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the instances within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as MIRSVM's overall superior performance against other multi-instance SVM models, having an average accuracy of $82.6\%$, which is $2.5\%$ better than the best performing state-of-the-art MI classifier.
\end{abstract}

@@

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as webmining and text categorization {\citep{Andrews2002}}, image classification and retreival \citep{qi2007,cao2017weakly}, human action recognition \citep{Yi2016}, medical diagnosis and imaging {\citep{Herrera2016}}, and bioinformatics {\citep{Dietterich1997}} (drug activity prediction and protein identification). In the case of drug activity prediction, MIL allows molecules to be represented as related instances contained in a bag rather than single instances with their own labels.

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a novel bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without making specific assumptions about the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

@@

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data~\citep{2017-KNOSYS-MIMV}. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. Our prososed contribution, MIRSVM, is an iterative method that uses SVMs to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities, performance, and sparser models.

@@

SimpleMI achieves the highest rank and competitive run times because, rather than using the instances in each bag to train a model, it takes the mean value of the instances in a bag and uses that for training. Even though SimpleMI has fast run-times, its performance over the previous metrics has been shown to be random and not as effective as the bag-level methods.

@@

The experimental study showed the better performance of MIRSVM compared with state-of-the-art multi-instance support vector machines, traditional multi-instance learners, as well as ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners, such as miGraph, Bagging, and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. This was especially reflected in the meta-ranks comparison, shown in Table~\ref{tab:metarank}, over Accuracy, Precision, Recall, Cohen's Kappa, AUC, and run time, where instance-level learners were ranked last, whereas bag-level and ensemble learners were better in rank, and MIRSVM ranked first. Our proposal, MIRSVM, performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.

MIRSVM could be further improved by invesitgating the following approaches. Rather than using a QP solver at each iteration of MIRSVM, a different optimal solver could be experimented with. Due to the run time implications of using a QP solver, using a different solver might improve MIRSVM's scalability. Another area to investigate would be implementing MIRSVM in high-performance distributed environments. A third concept to explore would be testing MIRSVM on highly imbalanced multi-instance datasets and observe its performance. MIRSVM could also be extended to handle multi-class multi-instance problems by using traditional methods such as one-versus-one, where $c(c - 1)/2$ classifiers are trained for a $c$-class problem, and one-versus-all, where a single classifier is trained per class~\citep{bishop2007pattern}.

>>>>

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the instances within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as MIRSVM's overall superior performance against other multi-instance SVM models{\color{red}, having an average accuracy of $82.6\%$, which is $2.5\%$ better than the best performing state-of-the-art MI classifier.}
\end{abstract}

@@

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as webmining and text categorization {\citep{Andrews2002}}, image classification and retreival \citep{qi2007,cao2017weakly}, human action recognition \citep{Yi2016}, medical diagnosis and imaging {\citep{Herrera2016}}, and bioinformatics {\citep{Dietterich1997}} (drug activity prediction and protein identification). {\color{red}In the case of drug activity prediction, MIL allows molecules to be represented as related instances contained in a bag rather than single instances with their own labels.}

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a novel bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without {\color{red}making specific assumptions about the bags' internal distributions}. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

@@

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data~\citep{2017-KNOSYS-MIMV}. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. {\color{red}Our prososed contribution,} MIRSVM, is an iterative method that uses SVMs to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities, performance, and sparser models.

@@

SimpleMI achieves the highest rank and competitive run times because, {\color{red}rather than using the instances} in each bag to train a model, it takes the mean value of the instances in a bag and uses that for training. Even though SimpleMI has fast run-times, its performance over the previous metrics has been shown to be random and not as effective as the bag-level methods.

@@

The experimental study showed the better performance of MIRSVM compared with state-of-the-art multi-instance support vector machines, traditional multi-instance learners, as well as ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners, such as miGraph, Bagging, and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. {\color{red}This was especially reflected in the meta-ranks comparison, shown in Table~\ref{tab:metarank}, over Accuracy, Precision, Recall, Cohen's Kappa, AUC, and run time, where instance-level learners were ranked last, whereas bag-level and ensemble learners were better in rank, and MIRSVM ranked first.} Our proposal, MIRSVM, performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.

{\color{red}MIRSVM could be further improved by invesitgating the following approaches. Rather than using a QP solver at each iteration of MIRSVM, a different optimal solver could be experimented with. Due to the run time implications of using a QP solver, using a different solver might improve MIRSVM's scalability. Another area to investigate would be implementing MIRSVM in high-performance distributed environments. A third concept to explore would be testing MIRSVM on highly imbalanced multi-instance datasets and observe its performance. MIRSVM could also be extended to handle multi-class multi-instance problems by using traditional methods such as one-versus-one, where $c(c - 1)/2$ classifiers are trained for a $c$-class problem, and one-versus-all, where a single classifier is trained per class~\citep{bishop2007pattern}. }
